Probability

### 2016-04-23 prob.tex
moved set valued to classical

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Probability Spaces}

%%%%%
\subsection{Classical Probability Spaces}

Probability
theory~\cite{inun.425605319950101,GrahamKnuthPatashnik1994,rohatgi2011introduction}
is defined using the notions of a \emph{sample space} $\Omega$, a
space of \emph{events}~$\events$, and a \emph{probability
  measure}~$\pmeas$. In this paper, we will only consider
\emph{finite} sample spaces: we therefore define a sample space
$\Omega$ as an arbitrary non-empty finite set, the space of events
$\events$ as $2^\Omega$, the powerset of $\Omega$, and the \emph{probability
measure} as a function $\pmeas : \events \rightarrow [0,1]$ such that:
\begin{itemize}
\item $\pmeas(\Omega) = 1$, and 
\item for a collection of pairwise disjoint events $E_i$, the probability
measures are additive $\pmeas(\bigcup E_i) = \sum \pmeas(E_i)$. 
\end{itemize}

\emph{Example of a problem on a finite sample space} (Two coin experiment)
Consider an experiment that tosses two coins. We have four possible outcomes that constitute the
  sample space $\Omega = \{ HH, HT, TH, TT \}$. The event that the
  first coin is ``heads'' is $\{ HH, HT \}$; the event that the two
  coins land on opposite sides is $\{ HT, TH \}$; the event that at
  least one coin is tails is $\{ HT, TH, TT\}$. Depending on the
  assumptions regarding the coins, we can define several probability
  measures. Here is a possible one:
\[\begin{array}{c@{\qquad\qquad}c}
\begin{array}{rcl}
\pmeas(\emptyset) &=& 0 \\
\pmeas(\{ HH \}) &=& 1/3 \\
\pmeas(\{ HT \}) &=& 0 \\
\pmeas(\{ TH \}) &=& 2/3 \\
\pmeas(\{ TT \}) &=& 0 \\
\pmeas(\{  HH, HT \}) &=& 1/3 \\
\pmeas(\{  HH, TH \}) &=& 1 \\
\pmeas(\{  HH , TT \}) &=& 1/3 
\end{array} & \begin{array}{rcl}
\pmeas(\{  HT, TH \}) &=& 2/3 \\
\pmeas(\{  HT , TT \}) &=& 0 \\
\pmeas(\{  TH , TT \}) &=& 2/3 \\
\pmeas(\{  HH, HT, TH \}) &=& 1 \\
\pmeas(\{  HH, HT, TT \}) &=& 1/3 \\
\pmeas(\{  HH, TH, TT \}) &=& 1 \\
\pmeas(\{  HT, TH, TT \}) &=& 2/3 \\
\pmeas(\{  HH, HT, TH, TT \}) &=& 1
\end{array}
\end{array}\]
Note that the probability measure for disjoint events such as $\{ HT \}$ and
$\{ TH \}$ do indeed add.

%%%%%
\subsection{Quantum Probability Spaces}

@@

In order to simplify the discussion, we will focus on $\mathscr{L}_{2}=\left\{ \textrm{impossible},\textrm{possible}\right\} $
where impossible and possible means $\left\{ 0\right\} $ and $\left(0,1\right]$
respectively while our result can be easily extend to other values
with the form $\left\{ \left\{ 0\right\} ,\left(0,\frac{1}{n}\right],\ldots,\left(\frac{n-1}{n},1\right]\right\} $.

Because $\mathscr{L}_{2}$ is a partition of $\left[0,1\right]$~\cite{Artin1991,Wechler1992,PurdomBrown1995,DummitFoote2004},
there is a natural surjective map~$\iota:\left[0,1\right]\rightarrow\mathscr{L}_{2}$
defined by $\iota\left(0\right)=\textrm{impossible}$ and $\iota\left(x\right)=\textrm{possible}$
for $x\ne0$. By applying $\iota$, we can define the corresponding
Born rule~$\bar{\mu}$ by $\bar{\mu}_{\psi}(E)=\iota\left(\mu_{\psi}(E)\right)$,
and the corresponding probability measure~$\hat{\mu}$ should satisfy: 
\begin{itemize}
\item $\hat{\mu}(\mathbb{1})=\textrm{possible}$, and 
\item for a collection of pairwise orthogonal $E_{i}$, we have $\hat{\mu}(\sum_{i}E_{i})=\bigvee_{i}\hat{\mu}(E_{i})$,
where $x\vee y=\textrm{impossible}$ if and only if $x=y=\textrm{impossible}$.
\end{itemize}
However, there is a $\mathscr{L}_{2}$-valued probability measure
\[
\hat{\mu}_{1}\left(E\right)=\begin{cases}
\textrm{impossible} & \textrm{, if }E=\proj{+};\\
\bar{\mu}_{0}(E) & \textrm{, otherwise.}
\end{cases}
\]
such that $\hat{\mu}_{1}\ne\bar{\mu}_{\psi}$ for all mixed state
$\ket{\psi}$.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Classical Probability Spaces}

A \emph{probability
  space}~\cite{inun.425605319950101,GrahamKnuthPatashnik1994,rohatgi2011introduction}
collects necessary conditions for reasoning coherently about
collections of uncertain events. It consists of a \emph{sample space}
$\Omega$, a space of \emph{events}~$\events$, and a \emph{probability
  measure}~$\pmeas$. In this paper, we will only consider
\emph{finite} sets of events: we therefore define a sample space
$\Omega$ as an arbitrary non-empty finite set and the space of events
$\events$ as $2^\Omega$, the powerset of $\Omega$. Given the set of
events $\events$, the \emph{probability measure} is a function
$\pmeas : \events \rightarrow [0,1]$ such that:
\begin{itemize}
\item $\pmeas(\Omega) = 1$, and 
\item for a collection $E_i$,  of pairwise disjoint events, 
  $\pmeas(\bigcup E_i) = \sum \pmeas(E_i)$.
\end{itemize}

\begin{example}[Two-coin probability space]
  Consider an experiment that tosses two coins. We have four possible
  outcomes that constitute the sample space
  $\Omega = \{ HH, HT, TH, TT \}$. The event that the first coin is
  ``heads'' is $\{ HH, HT \}$; the event that the two coins land on
  opposite sides is $\{ HT, TH \}$; the event that at least one coin
  is tails is $\{ HT, TH, TT\}$. Depending on the assumptions
  regarding the coins, we can define several probability
  measures. Here is a possible one:
\[\begin{array}{c@{\qquad\qquad}c}
\begin{array}{rcl}
\pmeas(\emptyset) &=& 0 \\
\pmeas(\{ HH \}) &=& 1/3 \\
\pmeas(\{ HT \}) &=& 0 \\
\pmeas(\{ TH \}) &=& 2/3 \\
\pmeas(\{ TT \}) &=& 0 \\
\pmeas(\{  HH, HT \}) &=& 1/3 \\
\pmeas(\{  HH, TH \}) &=& 1 \\
\pmeas(\{  HH , TT \}) &=& 1/3 
\end{array} & \begin{array}{rcl}
\pmeas(\{  HT, TH \}) &=& 2/3 \\
\pmeas(\{  HT , TT \}) &=& 0 \\
\pmeas(\{  TH , TT \}) &=& 2/3 \\
\pmeas(\{  HH, HT, TH \}) &=& 1 \\
\pmeas(\{  HH, HT, TT \}) &=& 1/3 \\
\pmeas(\{  HH, TH, TT \}) &=& 1 \\
\pmeas(\{  HT, TH, TT \}) &=& 2/3 \\
\pmeas(\{  HH, HT, TH, TT \}) &=& 1
\end{array}
\end{array}\]
The assignment satisfies the two constraints for probability measures:
the probability of the entire sample space is 1, and the probability
of every collection of disjoint events (e.g.,
$\{ HT \} \cup \{ TH \}$) is the sum of the individual probabilities.
\end{example}

In a strict computational or experimental setting, one may question
the reliance of the definition of probability space on the uncountable
and uncomputable real interval $[0,1]$. This interval includes numbers
like $0.h_1h_2h_3\ldots$ where $h_i$ is 1 or 0 depending on whether
Turing machine $M_i$ halts or not. Such numbers cannot be
computed. This interval also includes numbers like $\frac{\pi}{4}$
which can be computed but with increasingly large resources as the
precision increases. Therefore it is more appropriate in such
situations to consider probability measures mapping events to a finite
set of elements computable with a fixed set of
resources~\cite{Artin1991,Wechler1992,PurdomBrown1995,DummitFoote2004}. The
simplest such set, and the one we will consider exclusively in this
paper, is the set
$\mathscr{L}_{2}=\left\{
  \textrm{impossible},\textrm{possible}\right\}$
together with the operation $\vee$ where
$x \vee y=\textrm{impossible}$ if and only if
$x=y=\textrm{impossible}$. In relation to the first definition, one
can interpret impossible as the closed interval $[0]$, possible as the
open interval $(0,1]$, and $\vee$ as the union of intervals. The
definition of a probability measure in this case is modifed as being a
function $\pmeas : \events \rightarrow \mathscr{L}_{2}$ such that:
\begin{itemize}
\item $\pmeas(\Omega) = \textrm{possible}$, and 
\item for a collection $E_i$,  of pairwise disjoint events, $\pmeas(\bigcup E_i) = \bigvee \pmeas(E_i)$.
\end{itemize}

\begin{example}[Two-coin probability space with finite set-valued
  probability measure.] Under the new set-valued requirement, the
  probability measure in the first example becomes:
\[\begin{array}{c@{\qquad\qquad}c}
\begin{array}{rcl}
\pmeas(\emptyset) &=& \textrm{impossible} \\
\pmeas(\{ HH \}) &=& \textrm{possible} \\
\pmeas(\{ HT \}) &=& \textrm{impossible} \\
\pmeas(\{ TH \}) &=& \textrm{possible} \\
\pmeas(\{ TT \}) &=& \textrm{impossible} \\
\pmeas(\{  HH, HT \}) &=& \textrm{possible} \\
\pmeas(\{  HH, TH \}) &=& \textrm{possible} \\
\pmeas(\{  HH , TT \}) &=& \textrm{possible} \\
\end{array} & \begin{array}{rcl}
\pmeas(\{  HT, TH \}) &=& \textrm{possible} \\
\pmeas(\{  HT , TT \}) &=& \textrm{impossible} \\
\pmeas(\{  TH , TT \}) &=& \textrm{possible} \\
\pmeas(\{  HH, HT, TH \}) &=& \textrm{possible} \\
\pmeas(\{  HH, HT, TT \}) &=& \textrm{possible} \\
\pmeas(\{  HH, TH, TT \}) &=& \textrm{possible} \\
\pmeas(\{  HT, TH, TT \}) &=& \textrm{possible} \\
\pmeas(\{  HH, HT, TH, TT \}) &=& \textrm{possible} 
\end{array}
\end{array}\]
\end{example}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Quantum Probability Spaces}

@@

However, there is a $\mathscr{L}_{2}$-valued probability measure
\[
\hat{\mu}_{1}\left(E\right)=\begin{cases}
\textrm{impossible} & \textrm{, if }E=\proj{+};\\
\bar{\mu}_{0}(E) & \textrm{, otherwise.}
\end{cases}
\]
such that $\hat{\mu}_{1}\ne\bar{\mu}_{\psi}$ for all mixed state
$\ket{\psi}$.
